---
title: "STAT 327 Homework 3"
output: html_document
---

DUE FEBRUARY 9th

We'll grade your homework by

* opening your "hw3.Rmd" file in RStudio
* clicking "Knit HTML"
* reading the HTML output
* reading your "hw3.Rmd"

You should write R code anywhere you see an empty R code chunk. You
should write English text anywhere you see "..."; please surround it
with doubled asterisks (**...**) so that it will show up as boldface
and be easy for us to find.

Include reasonable labels (titles, axis labels, legends, etc.) with
each of your graphs.

Name: Tess Steplyk

Email: steplyk@wisc.edu

First, we need the `knitr` package. If loading it with `require()`
fails, then install it and load it. Then we load the package "MASS",
which contains the `Boston` data frame. (There's nothing for you to do
here.)

```{r,echo=FALSE}
if (!require("knitr")) {
  install.packages("knitr") # do this once per lifetime
  require("knitr") # do this once per session
}
require("MASS")
```

Run `?Boston` (outside this R Markdown document) to read the help page for the `Boston` data frame.

Convert the `chas` variable to a factor with labels "off" and "on" (referring to the Charles river).

```{r,echo=FALSE}
Boston$chas = factor(Boston$chas, levels = c(0, 1), labels = c("off", "on"))
```

How many rows are in the Boston data frame? How many columns?

```{r,echo=FALSE}
columns = ncol(Boston)
cat(sep = "", "The number of columns are ", columns, "\n")

rows = nrow(Boston)
cat(sep = "", "The number of rows are ", rows, "\n")
```

What does a row represent?
Each row in the data frame represents a unique suburb in the Boston area, containing statistics related to that suburb.

What does a column represent?
Each column in the data frame represents a different statistic that is being measured for each suburb.

Make a density plot (with rug) of tax rates.

```{r,echo=FALSE}
plot(density(Boston$tax), main = "Tax Rates")
rug(Boston$tax)
```

Describe the shape of the distribution of tax rates.
It's Bimodal around approximately 300 and 700

Note that the distribution shape doesn't make sense in light of the rug representation of the data. Make a histogram of the tax rates.

```{r,echo=FALSE}
hist(Boston$tax, breaks = 100, freq = TRUE, main = "Histogram of Tax Rates", 
    ylim = c(0, 140), xlim = c(100, 800))
```

Why is the second peak of the density plot so large? In what way is the rug representation of the data inadequate? Write a line or two of code to figure it out, and then explain it.

```{r,echo=FALSE}
max = length(which(Boston$tax == max(Boston$tax)))
cat(sep = "", "There are ", max, " occurrences of the maximum value, ", max(Boston$tax), 
    ", in this histogram.", "\n")

second = sort(Boston$tax, TRUE)[max + 1]
secondcount = length(which(Boston$tax == second))
cat(sep = "", "There are ", secondcount, " occurrences of the second highest value, ", 
    second, " ,in this histogram.", "\n")
```
There are multiple tax rates at 666 and 711, which is represented as a single dark line with the rug. It obscures the number of points at these tax rates.

Make a barplot of "chas".

```{r,echo=FALSE}
counts = table(Boston$chas)
barplot(counts, names.arg = NULL, main = "Charles River Count (chas)", ylim = c(0, 
    500))
```

How many neighborhoods are on the Charles river?

```{r,echo=FALSE}
Rivercount = length(which(Boston$chas == "on"))
cat(sep = "", "There are ", Rivercount, " neighborhoods on the Charles river.", 
    "\n")
```

Make a single graph consisting of three plots: 

* a scatterplot of "nox" on the y-axis vs. "dis" on the x-axis 
* a (vertical) boxplot of "nox" left of the scatterplot's y-axis
* a (horizontal) boxplot of "dis" below the scatterplot's x-axis

Hint: use `layout()` with a 4x4 matrix, using the top-right 3x3 corner
for the scatterplot, leaving the bottom-left 1x1 corner blank, and
using the other parts for the boxplots.

(An optional challenge, worth 0 extra credit points: remove the axis
and plot border from each boxplot.)

```{r, fig.width=7, fig.height=7}
m = matrix(data = c(1, 3, 3, 3, 1, 3, 3, 3, 1, 3, 3, 3, 0, 2, 2, 2), nrow = 4, 
    ncol = 4, byrow = TRUE)
layout(m)
boxplot(Boston$nox)
boxplot(Boston$dis, horizontal = TRUE)
plot(Boston$dis, Boston$nox)
```

